---
title: "European NATO Military Spending, 2014–2025"
author: "Edward McKelvey"
date: today
format:
  html:
    embed-resources: true
    toc: true
    code-fold: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(readxl)
source("R/sipri_functions.R")

sipri_gdp <- find_sipri_file() |>
  read_sipri() |>
  clean_sipri()

plot_data <- filter_countries(sipri_gdp)

# Values used in the text below
latest    <- max(plot_data$year)
val       <- function(ctry, yr) plot_data |> filter(country == ctry, year == yr) |> pull(share_gdp) |> round(1)
n_total   <- plot_data |> filter(year == latest) |> nrow()
n_2pc_now <- plot_data |> filter(year == latest, share_gdp >= 2) |> nrow()
n_2pc_14  <- plot_data |> filter(year == 2014,   share_gdp >= 2) |> nrow()
```

## Summary

In `r latest`, **`r n_2pc_now` of `r n_total`** European NATO members spent at least 2% of GDP on their militaries, up from `r n_2pc_14` in 2014. Poland stands out, reaching `r val("Poland", latest)`% of GDP, well above NATO's new 3.5% core target. Germany has roughly doubled its share since 2014, to `r val("Germany", latest)`%, while the UK is at `r val("United Kingdom", latest)`%.

## Spending since 2014

```{r chart}
#| fig-width: 9
#| fig-height: 5.5
plot_share_gdp(plot_data)
```

## Selected countries

```{r table}
plot_data |>
  filter(country %in% c("Poland", "Germany", "United Kingdom"),
         year %in% c(2014, 2022, latest)) |>
  select(country, year, share_gdp) |>
  pivot_wider(names_from = year, values_from = share_gdp) |>
  mutate(`Change since 2014 (pp)` = .data[[as.character(latest)]] - `2014`) |>
  rename(Country = country) |>
  knitr::kable(digits = 1, caption = "Military expenditure, % of GDP")
```

## Data and method

Data come from the [SIPRI Military Expenditure Database](https://www.sipri.org/databases/milex) (2026 release). The analysis covers the 29 European NATO members with armed forces; Iceland is excluded because it has none. SIPRI and NATO define military spending differently, so these figures will not always match NATO's official estimates.

All figures in this report are generated from code. See the [project README](README.md) for how to reproduce it.

*Personal project using publicly available data.*
